---
title: "2017-10-7 HW3_Q5"
author: "Ethan M. Lotz"
date: "October 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 5

5.) The following table consists of pixel level intensities for one spot patch area (these data are also available
in hmwk3.csv). In addition, those pixels outside the mask (circle/spot) have been labeled “O” while pixels in
the interior of the mask have been labeled “I”. Enter all data into the R programming environment (do NOT
use scan()). For these data, use the Mann-Whitney segmentation method to identify foreground and background pixels;
note the eight sampled background pixels have been identified for you in the third column. If you fail to
reject H0 let the number of pixels to discard equal 2. Once you have identified foreground pixels, calculate
the median intensity.

```{r}
Q5 <- read.csv("hmwk3.csv", header=TRUE)
y <- ifelse(Q5$Sample.background=="Yes", as.vector(Q5$Pixel.intensity), 0)
y <- y[! y %in% 0]

x <- ifelse(Q5$Sample.background!="Yes", as.vector(Q5$Pixel.intensity), 0)
x <- x[! x %in% 0]
x <- x[order(x)]
x
samplex <- x[1:8]
samplex
wilcox.test(samplex, y, alternative = "greater")
```

Fail to reject H0 so let the number of pixels to discard equal 2.
```{r}
samplex <- x[3:10]
samplex
wilcox.test(samplex, y, alternative = "greater")
```

Fail to reject H0 so let the number of pixels to discard equal 2.
```{r}
samplex <- x[5:12]
samplex
wilcox.test(samplex, y, alternative = "greater")
```

Fail to reject H0 so let the number of pixels to discard equal 2.
```{r}
samplex <- x[7:14]
samplex
wilcox.test(samplex, y, alternative = "greater")
```

```{r}
samplex <- x[9:16]
samplex
wilcox.test(samplex, y, alternative = "greater")
```

Reject H0 in favor of the alternative
```{r}
TrueForeground <- x[order(x)][9:length(x)]
TrueForeground

Median.Intensity <- median(TrueForeground)
Median.Intensity
```

